### 2015-03-25 swmpr_manu.tex
finished metabolism results and applications section

Water quality and weather observations from January 1995 to December 2014 for all \ac{NERRS} sites were obtained through a bulk data request using the \href{http://cdmo.baruch.sc.edu/aqs/zips.cfm}{zip downloads} feature of \ac{CDMO}.  After the download was complete, all csv files for each station were imported into R using the \texttt{import\_local} function, processed using the \texttt{setstep} and \texttt{qaqc} functions, then saved on a local hard drive as binary RData files (see \href{https://github.com/fawda123/swmp_rats/blob/master/R/dat_proc.R}{here} for the R script). This resulted in a single \texttt{swmpr} object for each paramter at each site.  All files were then uploaded to a remote server.  An \href{https://gist.github.com/fawda123/4fc51c2cb86341ed9291}{R script} was executed that retrieved and processed combined the water quality and weather data for each site.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/metab_plo}

@@

\caption[Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \ac{NERRS} reserve]{Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \ac{NERRS} reserve.  Values are daily integrated estimates as mean annual values averaged across all years with 95\% confidence intervals.  Two sites were chosen from each reserve that had the longest available time series.\label{fig:metab_plo}}
\end{figure}

@@

The second and third examples are two Shiny web applications \cite{Chang15} that illustrate summaries and comparisons of \ac{SWMP} data within and between reserves (see the \nameref{supp_info}).  The second application evaluates trends in \ac{SWMP} data within and between sites using an interactive map.  Trends between reserves can be viewed using the map, whereas trends at individual sites can be viewed by clicking on a map location.  Site-levels trends are described below the map using a simple linear regression to show an increase or decrease in values over time.  Trends on the map at each station are plotted as circles that identify the direction and significance of the trend.  The trend direction is blue for decreasing and red for increasing.  The significance is indicated by radius of the circle and color shading where larger points with darker colors indicate a strong trend.  The third application provides graphical summaries of water quality, weather, or nutrient station data at individual stations. The drop down menus can be used to select the station, date range, and parameter for plotting. Plots are based on daily averages for each parameter.  The data used for each application are similar to those used to estimate ecosystem metabolism described above.

>>>>

Water quality and weather observations from January 1995 to December 2014 for all \ac{NERRS} sites were obtained through a bulk data request using the \href{http://cdmo.baruch.sc.edu/aqs/zips.cfm}{zip downloads} feature of \ac{CDMO}.  All csv files for each station were imported into R using the \texttt{import\_local} function, processed using the \texttt{setstep} and \texttt{qaqc} functions, then saved on a local hard drive as binary RData files (see \href{https://github.com/fawda123/swmp_rats/blob/master/R/dat_proc.R}{here} for the R script). This resulted in a single \texttt{swmpr} object for each parameter at each site.  All files were then uploaded to a remote server.  An additional \href{https://gist.github.com/fawda123/4fc51c2cb86341ed9291}{R script} was executed that retrieved and processed combined the water quality and weather data for each site to esimate metabolism.  For brevity, only two sites at each reserve were evaluated.  Mean annual values at each site, organized by region, are shown in \Cref{fig:metab_plo}, whereas decadal comparisons are shown in \Cref{tab:metab_tab}.  All sites were generally net heterotrophic across the range of observations (i.e., sink of organic matter, in agreement with Caffrey \cite{Caffrey03}), although differences were observed in early (i.e., 1995-2004) as compared to recent (2005-2014) time periods. Overall, the results indicate that between-region and within-site differences in metabolism are apparent and varying by time period, such that a more comprehensive evaluation of factors that influence metabolic rates is needed.  More importantly, the use of the data retreival, synthesis, and analysis functions to create the results shown in \Cref{fig:metab_plo,tab:metab_tab} illustrates the utility provided by the SWMPr package.

% metabolism figure
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\textwidth]{figure/metab_plo}

@@

\caption[Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \ac{NERRS} reserve]{Aggregated estimates of net metabolism, gross production, and total respiration for two sites at each \ac{NERRS} reserve.  Values are daily integrated estimates as mean annual values averaged across all years with 95\% confidence intervals.  Two sites were chosen from each reserve that had the longest available time series. Regions are approximations of geographic coordinates and are not exact, e.g., Old Woman Creek (Lake Erie) is Atlantic.\label{fig:metab_plo}}
\end{figure}

@@

%latex.default(metabs_tab[, !names(metabs_tab) %in% c("site",     "region")], file = "", caption = cap, cgroup = c("NEM\\textsuperscript{{\\it a}}",     "Pg", "Rt"), n.cgroup = c(2, 2, 2), colheads = rep(c("1995-2004",     "2005-2014"), 3), rowlabel = "Site", rowname = metabs_tab$site,     digits = 2, insert.bottom = foots, rgroup = unique(metabs_tab$region),     n.rgroup = nrgroups, label = "tab:metab_tab", size = "footnotesize")%
\begin{table}[!tbp]
{\footnotesize
\caption{Trends in metabolism for two sites at each of the \ac{NERRS} reserves.  Values are averages of mean annual estimates for each period of observation (1994-2004 and 2005-2014).  Bold values indicate an increase from the first period, whereas italic values indicate a decrease.  Regions are approximations of geographic coordinates and are not exact, e.g., Old Woman Creek (Lake Erie) is Atlantic.\label{tab:metab_tab}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{2}{c}{\bfseries NEM\textsuperscript{{\it a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Pg}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Rt}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{1995-2004}&\multicolumn{1}{c}{2005-2014}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{1995-2004}&\multicolumn{1}{c}{2005-2014}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{1995-2004}&\multicolumn{1}{c}{2005-2014}\tabularnewline
\hline
{\bfseries Atlantic}&&&&&&&&\tabularnewline
~~acebb&-0.04&{\bf -0.03}&&5.11&{\it 3.81}&&-5.15&{\bf -3.82}\tabularnewline
~~acesp&-0.26&{\bf -0.1}&&4.68&{\bf 4.93}&&-4.94&{\it -5.03}\tabularnewline
~~cbmip&-0.14&{\it -0.17}&&1.02&{\bf 1.91}&&-1.16&{\it -2.08}\tabularnewline
~~cbmmc&-0.43&{\bf -0.24}&&6.06&{\bf 10.13}&&-6.48&{\it -10.38}\tabularnewline
~~cbvgi&0.13&{\it 0}&&6.56&{\it 4.47}&&-6.43&{\bf -4.46}\tabularnewline
~~cbvtc&-0.15&{\bf -0.07}&&5.58&{\bf 6.07}&&-5.73&{\it -6.14}\tabularnewline
~~delbl&-0.59&{\bf -0.16}&&6.68&{\it 6.51}&&-7.27&{\bf -6.67}\tabularnewline
~~delsl&-0.37&{\bf -0.33}&&2.84&{\bf 3.22}&&-3.21&{\it -3.55}\tabularnewline
~~grblr&&-0.02&&&1.28&&&-1.31\tabularnewline
~~grbor&&-0.04&&&3.16&&&-3.2\tabularnewline
~~gtmfm&-0.09&{\bf -0.04}&&3.36&{\bf 4.19}&&-3.45&{\it -4.23}\tabularnewline
~~gtmpc&-0.48&{\bf -0.45}&&1.95&{\bf 2.91}&&-2.43&{\it -3.37}\tabularnewline
~~hudsk&0.02&{\it 0.02}&&-0.36&{\it -0.39}&&0.38&{\bf 0.41}\tabularnewline
~~hudtn&-0.03&{\it -0.03}&&3.75&{\it 3.43}&&-3.78&{\bf -3.46}\tabularnewline
~~jacb6&0.05&{\it 0.03}&&3.41&{\bf 3.93}&&-3.36&{\it -3.89}\tabularnewline
~~jacb9&0.01&{\it -0.02}&&1.8&{\bf 3.1}&&-1.79&{\it -3.12}\tabularnewline
~~narnc&-0.64&{\bf -0.53}&&12.64&{\it 11.33}&&-13.25&{\bf -11.86}\tabularnewline
~~narpc&-0.14&{\bf -0.03}&&7.36&{\it 6.3}&&-7.5&{\bf -6.32}\tabularnewline
~~niwcb&-0.46&{\bf -0.28}&&5.5&{\bf 6.1}&&-5.95&{\it -6.38}\tabularnewline
~~niwdc&-0.13&{\it -0.2}&&7.28&{\bf 8.02}&&-7.41&{\it -8.23}\tabularnewline
~~nocec&-0.05&{\it -0.06}&&4.86&{\bf 4.88}&&-4.92&{\it -4.94}\tabularnewline
~~nocrc&-0.19&{\it -0.19}&&5.93&{\it 5.31}&&-6.12&{\bf -5.5}\tabularnewline
~~owcbr&-0.17&{\bf -0.09}&&1.45&{\bf 1.9}&&-1.62&{\it -1.99}\tabularnewline
~~owcwm&-0.36&{\bf -0.19}&&7.02&{\it 5.36}&&-7.38&{\bf -5.54}\tabularnewline
~~saphd&-1.28&{\bf -1}&&1.89&{\bf 3.77}&&-3.17&{\it -4.77}\tabularnewline
~~sapld&-0.16&{\it -0.43}&&2.32&{\bf 3.56}&&-2.49&{\it -3.99}\tabularnewline
~~welht&-0.07&{\bf -0.04}&&0.36&{\bf 0.63}&&-0.43&{\it -0.67}\tabularnewline
~~welin&-1.87&{\it -2.49}&&3.61&{\it 2.68}&&-5.48&{\bf -5.18}\tabularnewline
~~wqbcr&-0.01&{\bf 0.01}&&8.13&{\it 7.41}&&-8.14&{\bf -7.4}\tabularnewline
~~wqbmh&0.02&{\it 0}&&2.82&{\it 2.76}&&-2.79&{\bf -2.75}\tabularnewline
\hline
{\bfseries Gulf}&&&&&&&&\tabularnewline
~~apaeb&-0.35&{\it -0.35}&&4.19&{\bf 4.31}&&-4.54&{\it -4.67}\tabularnewline
~~apaes&-0.71&{\bf -0.67}&&3.35&{\it 3.05}&&-4.06&{\bf -3.73}\tabularnewline
~~gndbc&-1.02&{\bf -0.88}&&3.5&{\bf 4.3}&&-4.51&{\it -5.18}\tabularnewline
~~gndbh&-1.81&{\it -2}&&2.19&{\it 1.77}&&-4&{\bf -3.77}\tabularnewline
~~job09&-0.34&{\bf -0.25}&&5.32&{\bf 5.95}&&-5.66&{\it -6.2}\tabularnewline
~~job10&-0.15&{\it -0.15}&&3.58&{\bf 4.03}&&-3.72&{\it -4.2}\tabularnewline
~~marab&&-0.02&&&3.38&&&-3.4\tabularnewline
~~marce&&-0.01&&&3.62&&&-3.63\tabularnewline
~~rkbfb&-0.28&{\it -0.37}&&5.28&{\bf 6.25}&&-5.56&{\it -6.62}\tabularnewline
~~rkbfu&-0.33&{\it -0.41}&&2.95&{\bf 3.24}&&-3.28&{\it -3.64}\tabularnewline
~~wkbfr&-0.29&{\bf -0.14}&&6.27&{\bf 6.35}&&-6.57&{\bf -6.49}\tabularnewline
~~wkbwb&-0.29&{\bf -0.16}&&6.81&{\it 6.54}&&-7.1&{\bf -6.69}\tabularnewline
\hline
{\bfseries Pacific}&&&&&&&&\tabularnewline
~~elkap&0.03&{\bf 0.11}&&14.7&{\bf 19.84}&&-14.67&{\it -19.74}\tabularnewline
~~elknm&-0.24&{\bf -0.08}&&11.95&{\bf 16.37}&&-12.18&{\it -16.45}\tabularnewline
~~kachd&0.2&{\it -0.66}&&7.31&{\it 3.2}&&-7.12&{\bf -3.86}\tabularnewline
~~kacsd&-0.06&{\it -0.26}&&5.15&{\it 4.2}&&-5.21&{\bf -4.45}\tabularnewline
~~pdbbp&0.01&{\bf 0.08}&&6.29&{\bf 7.31}&&-6.27&{\it -7.23}\tabularnewline
~~pdbby&-0.07&{\bf -0.01}&&8.39&{\it 5.92}&&-8.47&{\bf -5.93}\tabularnewline
~~sfbcc&&-0.24&&&3.56&&&-3.79\tabularnewline
~~sfbgc&&-0.55&&&7.32&&&-7.87\tabularnewline
~~sosva&-0.2&{\bf -0.05}&&6.98&{\bf 7.83}&&-7.18&{\it -7.87}\tabularnewline
~~soswi&-0.28&{\bf -0.08}&&4.41&{\bf 4.74}&&-4.69&{\it -4.82}\tabularnewline
~~tjrbr&-1.33&{\bf -0.17}&&7.72&{\it 7.06}&&-9.05&{\bf -7.22}\tabularnewline
~~tjros&-0.26&{\it -0.34}&&10.18&{\bf 11.61}&&-10.44&{\it -11.96}\tabularnewline
\hline
\end{tabular}\end{center}}

\footnotesize{\textsuperscript{{\it a}}NEM: net ecosystem metabolism, Pg: gross production, Rt: total respiration, all values in g O$_2$ m$^{-2}$ d$^{-1}$ as annual averages.}\end{table}

The second and third examples are two Shiny web applications \cite{Chang15} that illustrate summaries and comparisons of \ac{SWMP} data within and between reserves that were created using the SWMPr package (see the \nameref{supp_info}).  The second application evaluates trends in \ac{SWMP} data within and between sites using an interactive map.  Trends between reserves can be viewed using the map, whereas trends at individual sites can be viewed by clicking on a map location.  Site-levels trends are described below the map using a simple linear regression to show an increase or decrease in values over time.  Trends on the map at each station are plotted as circles that identify the direction and significance of the trend.  The trend direction is blue for decreasing and red for increasing.  The significance is indicated by radius of the circle and color shading where larger points with darker colors indicate a strong trend.  The third application provides graphical summaries of water quality, weather, or nutrient station data at individual stations. The drop down menus can be used to select the station, date range, and parameter for plotting. Plots are based on daily averages for each parameter.  The data used for each application are similar to those used to estimate ecosystem metabolism described above.
